% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      


I would like to express my deepest gratitude to my mentor, Neil D. Lawrence for accepting me as a post graduate research student. Neil is a genuine role model, and while I have had many teachers, Neil’s mark is the most profound. Neil initiated me into the discipline of machine learning in graphical models and continuously taught me the most important scientific skills: how to dive deep into messy data and surface with simple models that address the question at hand, always striving to understand the connection between data, model and reality. Few have these skills and I was privileged to learn from a true master, I leave Neil with much yet to learn. Neil’s contribution to the research in this thesis is fundamental, from the basic idea of coupling Bayesian networks with gene networks to little comments that made my presentation so much clearer.

I have spent a total of ten terrific years as a student at the Hebrew University and this has been a significant chapter in my life. During this time, many teachers have molded me into the researcher I am today. I would like to thank Avi Wigderson for patiently teaching me the rigors of problem solving. Avi is a mental giant and I was most privileged to brainstorm with him and learn how he takes a hard problem apart into little bits he can understand. I would like to thank Shmuel Peleg for teaching me that resarch should first and foremost be fun. Shmuel taught me that if one does not find enjoyment and passion in the problem at hand, it is probably the wrong problem to be working on. I rarely left his office without a smile. I was especially fortunate to an adopting “mother” and “father”, Daphna Weinshall and Noam Nisan, in the Computer Science department. While never my official mentors, they took me under their wing, providing guidance, many rewarding discussions and emotional support. In addition, I would like to thank Noam for bringing to my attention α-modular functions and their connection to the MinReg algorithm. I would like to thank Daphna for actively fighting to make my years at the university more comfortable, be it easing the prerequisites when I transferred from mathematics or easing my TA workload as a new mother. 

Good science is always the joint effort of many people and the research in this thesis is no exception. This thesis could never have happened without Aviv Regev, my scientific partner, biology tutor and dearest friend. My research is the result of a close and synergistic collaboration with Aviv, working with whom is an absolute joy and pleasure. Aviv transformed me from a na ̈ıve computer scientist to semi-biologist teaching me so much more than biology along the way. In addition to sharing her wisdom and many unique insights, Aviv gave me endless support and backing. During the toughest and lowest points, Aviv was always there to stop me from quitting by infecting me with her energetic enthusiasm and leading me to believe in myself. There are simply no words to express my gratitude to her.

I would like to give many thanks to all my co-authors on the works presented here. Michal Linial, the first biologist who dared believe our ideas might have merit. Iftach Nachman, who shared with me the first steps of this research. Gal Elidan, who brought order and efficiency to the chaos in which I was used to be working in. It was a wonderful pleasure to work with Amos Tanay on our ‘underground’ MinReg project, and the speed in which he programmed some of our ideas never ceased to surprise me. Amos has great scientific vision and I cherish the many hours we spent brainstorming over coffee.

My intense collaboration with Eran Segal has been very fruitful and lead to great science. Eran, I very much admire your ability and stamina. I feel very priviliged to have worked with Daphne Koller, a brilliant scientist; I learned much from our many insightful discussions. Lots of thanks to all my lab mates at the Computational Biology group and the Machine Learning group at the Hebrew university. It was marvelous to belong to a group with such great academic cooperation and social atmosphere; Full of seminars, reading groups, or just hallway discussions; Beach parties, dinners, and hiking trips. Specifically, I would like to thank my office mate Matan Ninio, who fed me well, almost as often as he distracted me. Matan was always helpful from the countless times he aided me with system related issues, to the laborious work of printing this thesis and submitting it for me.

I would also like to thank the many people who gave me the support and technical backing so I could focus on my research. I thank the Ministry of Science, Israel, for the Eshkol fellowship awarded to me and the Higher Education Council, Israel, for additional financing. I thank the System
group at the Computer Science Department for the consistently providing the best and most reliable computer support possible. I thank the administrative staff at the Computer Science department for all their help and support, shielding me from the bureaucratic jungle that laid beyond our department. 

I would also like to thank Laura Garwin. Some times help comes unexpectedly, when my laptop crashed at critical stages of writing this thesis, Laura (at the time a stranger) out of pure kindness and generosity, lent me her personal laptop and hosted me in a wonderful office at the Bauer Center for Genomic Research.

During the course of my PhD. studies, the two most important events of my life occurred, the Births of Inbar and Carmel. I would like to thank my two most beloved daughters for distracting me and granting me joy and happiness of a magnitude I never knew before. I apologize to them, it is Inbar and Carmel that have paid the heaviest price for this thesis, during the endless hours I worked away from them. I hope you understand and forgive. I thank Rocha, my mother-in-law for the countless hours she took care of the girls, giving me more time to work. While Rocha was with
my girls, I could peacefully work, knowing they were getting the best of love and care. I thank my brother Michael for caring so much and for his constant reminders that there is so much more to life than research. I would like to thank Bat-Sheva for being available at any hour of the day or night for a relaxing walk and an opportunity to wind down.

I am grateful to both my parents, Mara and Aaron, for being such wonderful and supportive parents. They nurtured my curiosity, creativity and passion for understanding from the earliest age. I started my studies in the Mathematics department at the Hebrew university where both my parents
met and the completion of this thesis gives me a great feeling of fulfillment. Dad, thank you for attempting to teach me Cantor’s diagonal proof from preschool (that was a wee bit early), carefully correcting the English for this entire thesis and everything in between. Mom, you have and will always be my role model, you are my very inspiration to excel, I aspire to be like you.

Last, I dedicate this thesis to my better half, Itsik. I am endlessly indebted and grateful to Itsik for everything. My love, thank you for helping me in all aspects of my research. I thoroughly enjoyed our scientific discussions that occurred at all times of the day and in all forms of dress. Many of your comments have been invaluable to my work. Thank you for your help with all my manuscripts including this one. Thank you for your unconditional love in my worse moments and for being a strong pillar of support in most desperate moments. Thank you for making my victories
more memorable by sharing them with you, this victory could have never happened without all your encouragement and help.

My fellow PhD students helped me adjust to life in England, in particular to the, put it mildly, unfavourable local climate 


Moreover, I wish to thank the members of my committee at the Technische Universit ̈at Darmstadt, particularly my official “Doktorvater” Thomas Hofmann for his insightful comments on an early draft.


In the days before I submitted this thesis Jeremy Hill, Alexander Zien, Matthias Hein, Tobias Pfingsten, and especially Dilan G ̈or ̈ ur helped me a lot by proof-reading parts of (or even the whole) manuscript. Of course, any mistakes yet uncovered remain mine. Finally, this is also the place to thank my parents and their spouses for their generous support and encouragement during my studies and before.
The most beautiful thing during the past years was to have Claudia Beckmann by my side. Doing research and writing a thesis involves moments of doubt and frustration. But she always knew how to cheer me up and has proven an enormous patience with me in times when I lost the necessary distance to my work.


\end{acknowledgements}
