% Appendix A

\chapter{Summery of DDP Progress} % Main appendix title

\label{AppendixB} % For referencing this appendix elsewhere, use \ref{AppendixA}

\lhead{Appendix A. \emph{Summery of DDP Progress}} % This is for the header on each page - perhaps a shortened title

Based on the plan of the Doctoral Development Programme (DDP) various activities have been completed-

\section{Posters Presentation and talks}
\subsection{Posters}
\begin{itemize}
 \item Rahman M A,  Lawrence N D, ``A Probabilistic Dynamic Model for Transcription Factor Activity of C. Elegans'', Machine Learning Summer School and AISTATS joint poster session, Reykjavic, Iceland 2014.
\end{itemize}
\subsection{Workshop Notebook}
\begin{itemize}
 \item Rahman M A,  Lawrence N D, ``Inferring Transcription Factor Activities by Combining Binding Information with Gene Expression Profiles'', Available online, url-
 \url{http://nbviewer.ipython.org/github/SheffieldML/notebook/blob/master/compbio/transcriptionFactorActivities.ipynb}
\end{itemize}
\subsection{Talks}
\begin{itemize}
 \item Rahman M A, Lawrence N D, ``Dynamic model for transcription factor activity of C. Elegans'', Sheffield Institute for Translational Neuroscience (SITraN), University of Sheffield, November 2013.
 \item Rahman M A, Lawrence N D, ''A probabilistic dynamic model for transcription factor activity of C. Elegans``, Sheffield Institute for Translational Neuroscience (SITraN), University of Sheffield, May 2013.
\end{itemize}


\section{Participation in Conference, Workshop and Summer schools}

As a part of DDP progress I have participated a number of Conference, Workshop and Summer schools given bellow-
\begin{itemize}
\item Gaussian Process Summer School 2013, University of Sheffield, UK was organized by our lab ML@SITraN from June 10 to June 12, 2013 followed by Workshop on Latent Force Model 2013. I was an organizing member here.
\item I have attended the Natural Computing Applications Forum (NCAF) Meeting at University of Oxford, Oxford, UK from July 04 to July 05, 2013.
\item I have attend the Machine Learning Summer School (Advanced Topics in Machine Learning) at Technical University of Denmark, Copenhagen, Denmark on August 2013.
\item Gaussian Process Winter School 2014, University of Sheffield, UK was organized by our lab ML@SITraN from January 13 to January 16, 2014. I was a organizing member here. Workshop on Spatiotemporal Modelling with Gaussian Processes was collocated with GPWS 2014.
\item I have attended the Mathematical and Statistical Aspects of Molecular Biology (MASAMB) 2014 workshop at University of Sheffield, UK from April 09 to April 10, 2014
\item I have attended the conference of Artificial Intelligence and Statistics (AISTATS) 2014 at Reykjavik, Iceland followed by the Machine Learning Summer School (MLSS2014).
\item Gaussian Process Summer School 2014, University of Sheffield, UK was organized by our lab ML@SITraN from September 15 to September 17, 2014. Workshop on Gaussian Processes for Feature Extraction was collocated with GPSS 2014. I was a organizing member here.
\end{itemize}

\section{Demonstration}
As a part of DDP progress the activities relating to teaching and demonstrating are given bellow-
\begin{itemize}
\item COM4509/COM6509 Machine Learning and Adaptive Intelligence with Professor Neil D Lawrence(ACADEMIC YEAR 2014-15). 
\item COM161: Introduction to Programming and Problem-Solving and COM160: Computer Problem Solving and Object-Oriented Design with Professor Mark Hepple (ACADEMIC YEAR 2014-15).
\item COM1005/COM2007: Machines and Intelligence with Professor Rob Gaizauskas (ACADEMIC YEAR 2014-15).
\item FCE101: Introduction to Bioengineering (ACADEMIC YEAR 2011-12) with Professor Neil D lawrence.
\end{itemize}

\section{Participation in DDP Courses}
To improve the core concepts of research I have attended the following courses/modules:

\begin{itemize}
 \item COM - COM4509~COM6509: Machine Learning and Adaptive Intelligence(GRADUATE YEAR 2011-12).
 \item COM4508/6508: Introduction to Computational Systems Biology (GRADUATE YEAR 2011-12).
 \item FCE - FCE6100: Research Ethics and Integrity (GRADUATE YEAR 2011-12).
 \item COM - COM6930: Research Teamwork in Computer Science (GRADUATE YEAR 2011-12).
 \item COM - COM6940: Presentation of Computer Science Research (GRADUATE YEAR 2011-12).
 \item COM - COM6950: Developments in Computer Science (GRADUATE YEAR 2011-12).
\end{itemize}

\section{Other activities related to DDP}
\begin{itemize}
 \item Take part in the poster presentation at the research retreat of Department of Computer Science, University of Sheffield.
 \item Reading, understanding and writing about relevant research, and implementing simple algorithms to improved research related skills.
 \item Maintaining a personal website detailing any research and experiments that have been done.
 \item Made contact with experts from a different department/universities to get the in-depth knowledge on Machine Learning and System Biology.
\end{itemize}